---
title: "Data Wells: Race and State Violence in the United States from 1892"
subtitle: "Quantitative Histories Workshop"
author:
  - name: Nathan Alexander
    affiliations:
      - ref: hu
  - name: Kade Davis
    affiliations:
      - ref: mc
  - name: Basil Ghali
    affiliations:
      - ref: hu
  - name: Qyana Stewart
    affiliations:
      - ref: hu
  - name: Gabriella La Cour
    affiliations:
      - ref: sc
affiliations:
  - id: hu
    name: Howard University
    city: Washington
    state: DC
  - id: mc
    name: Morehouse College
    city: Atlanta
    state: GA
  - id: sc
    name: Spelman College
    city: Atlanta
    state: GA
  - id: cd
    name: Center for Applied Data Science and Analytics
    city: Washington
    state: DC
format:
  pdf:
    pdf:
    toc: true
    number-sections: true
    colorlinks: true
    self-contained: true
    embed-resources: true 
  html:
    self-contained: true
    embed-resources: true
  revealjs:
    output-file: 2024_04_27_bob_moses.qmd
    height: 900
    width: 1600
    smaller: false
    scrollable: true
    slide-number: c/t #< collapsted/total
    footer: ""
    logo: "howard-logo.png"
    toc: false
    echo: true
    incremental: false
    self-contained: true
    theme: default
---

## [*Data*]{style="color:darkgreen;"} + Ida B. [*Wells*]{style="color:darkgreen;"}-Barnett

::: {.incremental}

  - We use the term "[Data Wells]{style="color:darkgreen;"}" to describe how we practice the identification, input, and storage of what can be termed as critical insights data, or CIDs.
  
  - We use information in databases in four ways: 
  
    (1) studying problems in the [quantification of historical information]{style="color:red;"} across various axes: time, social constructs, and/or systemic issues, 
    
    (2) data [identification]{style="color:red;"} and [wrangling]{style="color:red;"}, 
    
    (3) data [analysis]{style="color:red;"} and [communication]{style="color:red;"}, and 
    
    (4) [modeling]{style="color:red;"} abstract inquiries.
  
  - We begin our analysis with Ida B. Wells-Barnett's organization and analysis of lynching.
  
  - We describe "Data Wells" of U.S. state violence using quantitative history as a frame.

:::

# Quantitative Histories Workshop {background-color="lightgrey"}

`Curriculum & software development collective` 

and

[research lab]{style="color:orange;"}

```{r}
#| echo: false
#| output: false
#| warning: false

# install all required packages
require("terra")
require(tidyverse)
install.packages("tidycensus", repos = "http://cran.us.r-project.org")
require(tmap)
require("ggiraph")
require(mapdeck)
require(ggplot2)

# load all required libraries
library(tidycensus)
library(tidyverse)
library(tigris)
library(tidycensus)
library(tidyverse)
library(patchwork)
library(scales)
library(ggplot2)
library(sf)
library(patchwork)
```

## Quantitative history

:::: {.columns}

::: {.column width="40%"}

  - Quantitative history considers methods and approaches to artifacts as data and information.
  
  - Historians like Pierre Chanu (book to the right) are centered in traditional texts; more traditional perspectives are uncovering troubling practices with regard to race[^chanu].
  
  - Despite many long-standing critiques, there are [fewer critical dimensions]{style="color:orange;"} in [quantitative history] narratives.
  
:::

[^chanu]: Vovelle, Michel (1987). Bourgeoisies de province et Revolution. Presses Universitaires de Grenoble.

::: {.column width="60%"}

![[*Histoire Quantitative*](https://www.jstor.org/stable/1876930){target="blank"} by Pierre Chanu](quanthist.jpg){width=40%}

:::

::::

## Racialization and U.S. State Violence

Today, we will explore three forms of state sponsored violence.

. . .

  - Lynching

. . .

  - Police

. . .

  - Prisons

## The Red Record

  - **Personal experience**. In 1892, a close friend of Ida B. Wells-Barnett was lynched. Wells-Barnett, a known activist, community organizer, and journalist, would generate quantitative indicators of lynching as state violence.

. . .

  - **Intuition and method**. Like many Black communities at the time and other allies, Wells-Barnett acknowledged both the personal (micro) social forces of racism and the systemic (macro) nature of white racial violence. In this case, this violence was expressed through the practice of lynching.

. . .

  - **Impact**. Wells-Barnett's databases, and the use of number and quantification have a profound impact on the current view of state-sponsored racial violence.

## Lynching

Caitlin Pollock has created [software](https://redrecord.cmjpollock.com/) based on a series of extracted data from Wells-Barnett's work. Although the data provides for quick loading and analysis, it does require some data wrangling.

::: {.panel-tabset}

### 1893

Content for 1893

```{r}
#| echo: false
#| output: true
#| warning: false
df1893 <- read_sf("https://raw.githubusercontent.com/quant-shop/IdaBWellsProject/master/RedRecord/redrecord1893.geojson.txt")
df1893

```

### 1894

Content for 1894

```{r}
#| echo: false
#| output: true
#| warning: false
df1894 <- read_sf("https://raw.githubusercontent.com/quant-shop/IdaBWellsProject/master/RedRecord/redrecord1894.geojson.txt")
df1894

```

### 1895

Content for 1895

```{r}
#| echo: false
#| output: false
#| warning: false
install.packages("xml2", repos = "http://cran.us.r-project.org")
install.packages("XML", repos = "http://cran.us.r-project.org")
library(xml2)
library(XML)
df1895t <- read_xml("https://raw.githubusercontent.com/quant-shop/IdaBWellsProject/master/RedRecord/redrecord1895.xml")
df1895t
df1895s <- xmlParse(df1895t)
```

```{r}
#| echo: false
#| output: true
#| warning: false
df1895s
```

:::

---

## Concerns

Pollock deals with the issue of erasure in their development of the data.

## Police


## Prisons

::: {.panel-tabset}

### Federal

Content for federal

### State

Content for state

### Federal and State

Content for federal and state

:::


# Thank you
. . .

Thank you for joining us and citing today's presentation.

Alexander, N., Davis, K., Ghali, B., Stewart, K., & La Cour, G. (2024, April 26). Data Wells: Race and State Violence in the United States from 1892. *The 2024 Bob Moses Conference*. [Online](https://www.bobmosesconference.com/){target="blank"}.

